Face recgontion using Particle Swarm clustering

### 2011-08-13 FacePaper/paper.tex
adding most from linux

%The Nelder-Mead smplex method makes use of a construct called a simplex (see
%Figure 1.). When the search space is n-dimensional, the simplex consists of n+1 solutions, s,, i = {1, 2,\dots  , n+1}, that are usually closely spaced. As shown in the top left of Figure 1., in a two-dimensional search plane, a simplex is a triangle. The fitness of each solution is considered in each step of the Nelder-Mead method, and the worst solution w is identified. The centroid, c, of the remaining n points is computed and the reflection of walong it determined. This reflection yields a new solution r that replaces w, in the next step, as shown in the top right of Figure 1. If the solution r produced by this reflection has a higher fitness than any other solution in the simplex, the simplex is further expanded along the direction of r, as shown in the middle left of the figure. On the other hand, if r has a low fitness compared to the others, the simplex is contracted. Contraction can be either outward or inward depending upon whether r is better or worse than w. The contraction operations are shown in the middle right and bottom left of the figure. If neither contraction improves the worst solution in the simplex, the best point in the simplex is computed, and a collapse is then carried out, and all the points of the simplex are moved a little closer towards the best one, as shown in the bottom right of the same figure.

@@

After the computing the eigenfaces of every person on the database the system
extracts eigen weight vector for each image. Equation \ref{} shows how the
weight vector is computed from the eigen faces. These weights are used as
features for the clustering algorithm.

\subsection{PSO clustering}
The clustering algorithm is used to train the system to identify the faces in
the dataset. The clustering algorithm used is based on Particle algorith based
on the Merwe and Engelbrechet \cite{psoclustering} Method. Each particle
$X_i$ is constructed such that that
\begin{equation}
X_i=\left( m_{i1}, m_{i2}, m_{i3},\dots, m_{iN_c}\right)
\end{equation}
where $N_c$ is the number of clusters to be formed and $m_{ij}$ corresponds to the $j^{th}$ centroid of the  $i^{th}$ particle, the centroid of the cluster $C_{ij}$. Thus single particle represents a candidate solution to a given clustering algorithm. Each particle is evaluated using the following equation:
% \begin{equation}
% F_c=\frac{\sum^{N_c}_{j=1}\right[\sum_{\forall Z_j \epsilon C_{ij}}\frac{ d\right(Z_p,m_{ij}\left)}{|C_{ij}|}  \left]}{N_c}
% \end{equation}
\begin{equation}
F_c  = \frac{{\sum\nolimits_{j = 1}^{N_c } {\left[ {\sum\limits_{\forall z_j  \in C_{ij} } {\frac{{d(z_p ,m_{ij} )}}{{\left| {C_{ij} } \right|}}} } \right]} }}{{N_c }}
\end{equation}
where $F_c$ is the fitness function, $N_c$ is the number of clusters in the particle, $Z_p$ denotes the $p^{th}$ data vector $\left| {C_{ij} } \right|$ is the number of data vectors belonging to the cluster $C_{ij}$ and $d$ is the Euclidian distance between $Z_p$ and $m_{ij}$.

@@

\begin{figure}
 \begin{center}
 \centering
%\includegraphics[scale=0.5]{blockchart.jpg}
 \subfigure['PSO
 clustering ']{\label{fig:pso1}\includegraphics[width=0.25\textwidth]{pso1.jpg}}
 \subfigure['Nelder
 meed a']{\label{fig:pso2}\includegraphics[width=0.25\textwidth]{pso2.jpg}}
  \subfigure['Nelder
  meed b']{\label{fig:pso3}\includegraphics[width=0.25\textwidth]{pso3.jpg}}
\caption{ Particle Clustering Alogirhtm}
\label{fig:PSOswarm}
 \end{center}\end{figure}
Both previous algorithms  were implemented and tested for correctness. Also the k-means algorithm was used to compare the result of the PSO clustering algorithm. The code was implemented using Matlab and using a Database as train and test set. The Databases contains 50 persons and 4 images per person. Currenlty all the images are used for training and the same set is used for testing.

The PSO algorithm is used to clusters each face weights computed using equation \ref{weighteq}. The algorithms first compute the eigen vectors for all 50 faces as explained in section \ref{facereg}. The computed $u_j$ is used to compute the weights for each image in the database is computed which generate a data matrix of 150 X 50 ( number of image X weight vector (no of eigenfaces)). The clustering algorithm is used to get center weigh vector for each person. I choose the number of clusters same as number of person for easier recognition step.

@@

\subsection{Experiments Finding PSO paramters}
\begin{table}
	\centering
		\caption{PSO Result }
	\label{tab:PSO1}
	%\scalebox{0.99}{
		\begin{tabular}{|l|c|c|c|c|c|}
		 \hline
C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
W	&0.5&	1	&1.5	&2&	4	\\ \hline
	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
\end{tabular}
		%}
\end{table}

\subsection {Comparing With other systems}

\subsection{ Comparing Nelder enhancement}

\section{Conclusion}
\label{sec:Conclusion}
\section{future work}
% \subsection{Subtitle}
%
% Plain text.
%
% % \subsection{Another subtitle}
%
% More plain text.
\bibliographystyle{model1a-num-names}
\bibliography{mybibfile,library}

>>>>

The Nelder-Mead smplex method makes use of a construct called a simplex When the search space is $n$-dimensional, the simplex consists of $n+1$ solutions, $s i = {1,2,\dots  , n+1}$, that are usually closely spaced. In a two-dimensional search plane, a simplex is a triangle. The fitness of each solution is considered in each step of the Nelder-Mead method, and the worst solution $w$ is identified. The centroid, $c$, of the remaining n points is computed and the reflection of $w$ along it determined. This reflection yields a new solution $r$ that replaces $w$, in the next step. If the solution $r$ produced by this reflection has a higher fitness than any other solution in the simplex, the simplex is further expanded along the direction of $r$. On the other hand, if $r$ has a low fitness compared to the others, the simplex is contracted. Contraction can be either outward or inward depending upon whether $r$ is better or worse than $w$. The contraction operations are shown in the middle right and bottom left of the figure. If neither contraction improves the worst solution in the simplex, the best point in the simplex is computed, and a collapse is then carried out, and all the points of the simplex are moved a little closer towards the best one.

@@

After the computing the eigenfaces of every person on the database the system extracts eigen weight vector for each image. Equation \ref{} shows how the weight vector is computed from the eigen faces. These weights are used as features for the clustering algorithm. The output of this stage is a dataset that contains a eigen weight vector for each face in the database which is considered the feature vector of this face.

\subsection{PSO clustering}

The clustering algorithm is used to group the input faces into clusters which represents the persons in the dataset. The process of clustring the input faces is simillar to a classifier training in classification systems.    The clustering algorithm used is based on Particle Swarm algorithm  based on the Merwe and Engelbrechet \cite{psoclustering} Method where particle represents a possible clustering of the whole dataset.  Each particle $X_i$ is constructed such that that
\begin{equation}
X_i=\left( m_{i1}, m_{i2}, m_{i3},\dots, m_{iN_c}\right)
\end{equation}
where $N_c$ is the number of clusters to be formed and $m_{ij}$ corresponds to the $j^{th}$ centroid of the  $i^{th}$ particle, the centroid of the cluster $C_{ij}$. Thus single particle represents a candidate solution to a given clustering algorithm. Each particle is evaluated using the following equation:
% \begin{equation}
% F_c=\frac{\sum^{N_c}_{j=1}\right[\sum_{\forall Z_j \epsilon C_{ij}}\frac{ d\right(Z_p,m_{ij}\left)}{|C_{ij}|}  \left]}{N_c}
% \end{equation}
\begin{equation}
F_c  = \frac{{\sum\nolimits_{j = 1}^{N_c } {\left[ {\sum\limits_{\forall z_j  \in C_{ij} } {\frac{{d(z_p ,m_{ij} )}}{{\left| {C_{ij} } \right|}}} } \right]} }}{{N_c }}
\end{equation}
where $F_c$ is the fitness function, $N_c$ is the number of clusters in the particle, $Z_p$ denotes the $p^{th}$ data vector $\left| {C_{ij} } \right|$ is the number of data vectors belonging to the cluster $C_{ij}$ and $d$ is the Euclidian distance between $Z_p$ and $m_{ij}$.

@@

\begin{figure}
 \begin{center}
 \centering
%\includegraphics[scale=0.5]{blockchart.jpg}
 \subfigure['PSO
 clustering ']{\label{fig:pso1}\includegraphics[width=0.25\textwidth]{pso1.jpg}}
 \subfigure['Nelder
 meed a']{\label{fig:pso2}\includegraphics[width=0.25\textwidth]{pso2.jpg}}
  \subfigure['Nelder
  meed b']{\label{fig:pso3}\includegraphics[width=0.25\textwidth]{pso3.jpg}}
\caption{ Particle Clustering Alogirhtm}
\label{fig:PSOswarm}
 \end{center}\end{figure}
%Both previous algorithms  were implemented and tested for correctness. Also the k-means algorithm was used to compare the result of the PSO clustering algorithm. The code was implemented using Matlab and using a Database as train and test set. The Databases contains 50 persons and 4 images per person. Currenlty all the images are used for training and the same set is used for testing.

The PSO algorithm is used to clusters each face weights computed using equation \ref{weighteq}. The algorithms first compute the eigen vectors for all faces in the database as explained in section \ref{facereg}. The computed $u_j$ Eigen vectors is used to compute the weights for each input image in the database. The clustering algorithm is used to get center weigh vector for each person. %I choose the number of clusters same as number of person for easier recognition step.

Integrating the PSO clustering algorithms with the Nelder Mead algorithm is done using either using Nelder-Mead in a parallel iterative process with the PSO as in figure \ref{fig:pso2}. Figure \ref{fig:pso3} shows a sequencial integration of Nelder-Mead search method inside the PSO algorithm where after each iteration of the PSO a single interation of Nelder-Mead is performed to improve the solution. Both mehtods are embeded in the system and comparison  between both of them is presented in \ref{sec:Results}, % After the PSO clustering algorithm is constructed two d

@@

\subsection{Experiments Finding PSO paramters}
\label{sec:pso}

Particle swarm consist of many paramters that controls how the search is operated and have direct impact on the output of the system.

\begin{table}
	\centering
		\caption{PSO Result }
	\label{tab:PSO1}
	%\scalebox{0.99}{
		\begin{tabular}{|l|c|c|c|c|c|}
		 \hline
C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
W	&0.5&	1	&1.5	&2&	4	\\ \hline
	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
\end{tabular}
		%}
\end{table}

\subsection {Comparing With other systems}
\label{sec:comparingSystems}

\begin{table}
	\centering
		\caption{PSO Result }
	\label{tab:system}
	%\scalebox{0.99}{
		\begin{tabular}{|l|c|c|c|c|c|}
		 \hline
C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
W	&0.5&	1	&1.5	&2&	4	\\ \hline
	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
\end{tabular}
		%}
\end{table}

\subsection{ Comparing Nelder enhancement}
\label{sec:comparingSystems}
Figrue \ref{} shows two different method to add Nelder enhancment to the PSO clustering algorithm. As mentioned in section \ref{} both methods are evaluated to select best integration methods.  Table \ref{tab:nelder} shows the result of comparison between the clustering algorithm without Nelder-Mead and using two integration methods. The results show that Nelder-Mead enhancment improved the recognition accuracy. %achieve

\begin{table}
	\centering
		\caption{PSO Result }
	\label{tab:nelder}
	%\scalebox{0.99}{
		\begin{tabular}{|l|c|c|c|c|c|}
		 \hline
C1 &	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.893	&70.893	&70.605	&70.029	&69.741 \\  \hline
C2	&	1	&1.2&	1.6&	1.8 &	2	\\ \hline
	&70.89	&70.61	&70.89	&69.45	&69.45 \\ \hline
W	&0.5&	1	&1.5	&2&	4	\\ \hline
	&69.74	&70.89&	70.89	&70.32&	70.89	\\ \hline
\end{tabular}
		%}
\end{table}

\section{Conclusion and future work}
\label{sec:Conclusion}
In this paper we introduced a system that recognize faces using a hybrid PSO clustering algorithm. The clustering algorithm is simple and achieved better result than other methods.

Our future work will focus on improving the feature extraction method by using various features like DCT and wavelet features. Another improvement is trying to use clustering algorith to problems rather than face recognition.

%\section{future work}
% \subsection{Subtitle}
%
% Plain text.
%
% % \subsection{Another subtitle}
%
% More plain text.
\bibliographystyle{model1a-num-names}
\bibliography{mybibfile,library}
